--- 
title: "Heavy Metal Analysis Notebook"
author: "Todd Kelman"
date: "2/12/2019"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(googlesheets)
library(readxl)
library(janitor)
```


```{r sheet processing, include = FALSE}
# open Pb/PBDE/PCB Egg Submissions using the direct key since googlesheets package can't
# use gs_title (or gs_ls()) for Team Drives:
sheet_meta <- gs_key("1kgXukJLuWqII972QHkFm5o8JutPybPZZ-_UTRa2oOfk", lookup = FALSE, visibility = "private")
sheet_data <- gs_read(sheet_meta, ws = "HVM Values Only", col_names = TRUE, skip = 0, verbose = TRUE, literal = FALSE)
premise_directory <- read_excel("Data/Premise_Directory 9-4-18 geocoded.xls")

cn_sheet_data <- clean_names(sheet_data)
cn_premise_directory <- clean_names(premise_directory)
```

```{r extact_data, , include = FALSE}
# extract the metals of interest
metal_subset <- cn_sheet_data %>% 
  select(bp_id, sample_id, zip, county, matches("57g")) %>%
  mutate(zip = str_sub(as.character(zip), 1, 5)) %>%      # convert from zip+5 to just zip
  mutate(county = str_to_upper(county)) %>%
  mutate(premise_id = str_extract(sample_id, "[^-]+")) 
```

## Assessing variance

The first step is assessing the variability in lead concentrations in eggs submitted
from the same premise.  The following graph is for non-commercial premises submitting
at least 5 eggs, with each lead value represented as a dot superimposed on a
standard boxplot showing median and quartile values

```{r variance_5plus_eggs}
metal_subset %>% 
  add_count(premise_id) %>% 
  filter(n>4) %>% 
  filter(!str_detect(tolower(premise_id), pattern = "c")) %>% 
  mutate(premise_id = as.integer(premise_id)) %>% 
  arrange(premise_id) %>% 
  ggplot(mapping = aes(x = factor(premise_id), y = pb_ug_57g)) +
    geom_boxplot() +
    geom_dotplot(binaxis='y', 
                stackdir='center', 
                dotsize = 5, 
                binwidth = .02,
                fill="red") +
    labs(x = "Premise ID", y = "ug Pb per 57g egg", 
         title = "Lead values for noncommercial premises with at least 5 eggs submitted")
```

So it looks like about half the samples are exhibiting high variability, which 
has implications for pooling and how we should evaluate the non-pooled vs. pooled
data.  
  
Here is the identical analysis for premises with two to four eggs:  

```{r variance_2to4_eggs}
metal_subset %>% 
  add_count(premise_id) %>% 
  filter(n>1, n<5) %>% 
  filter(!str_detect(tolower(premise_id), pattern = "c")) %>% 
  mutate(premise_id = as.integer(premise_id)) %>% 
  arrange(premise_id) %>% 
  ggplot(mapping = aes(x = factor(premise_id), y = pb_ug_57g)) +
  geom_boxplot() +
  geom_dotplot(binaxis='y', 
               stackdir='center', 
               dotsize = 3, 
               binwidth = .02,
               fill="red") +
  labs(x = "Premise ID", y = "ug Pb per 57g egg", 
       title = "Lead values for noncommercial premises with 2-4 eggs submitted")
```

...which shows a very similar pattern:  fairly high variability between submissions.
