---
title: "Heavy Metal Analysis"
author: "Todd Kelman"
date: "2/12/2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(googlesheets)
library(readxl)
library(janitor)
```


```{r sheet processing, include = FALSE}
# open Pb/PBDE/PCB Egg Submissions using the direct key since googlesheets package can't
# use gs_title (or gs_ls()) for Team Drives:
sheet_meta <- gs_key("16iR75UiNn_PPle-ORK9sVbWFUJPlVIuHBhQlEJ9Ayuk", lookup = FALSE, visibility = "private")
sheet_data <- gs_read(sheet_meta, ws = "HVM Results", col_names = TRUE, skip = 0, verbose = TRUE, literal = FALSE)
premise_directory <- read_excel("Data/Premise_Directory 9-4-18 geocoded.xls")

cn_sheet_data <- clean_names(sheet_data)
cn_premise_directory <- clean_names(premise_directory)
```

```{r extact_data}
# extract the metals of interest
metal_subset <- cn_sheet_data %>% 
  select(bp_id, sample_id, zip, county, matches("57g")) %>%
  mutate(zip = str_sub(as.character(zip), 1, 5)) %>%      # convert from zip+5 to just zip
  mutate(county = str_to_upper(county)) %>%
  mutate(premise_id = str_extract(sample_id, "[^-]+")) 
```

## Assessing variance

The first step is assessing the variability in lead concentrations in eggs submitted
from the same premise.  The following graph is for non-commercial premises submitting
at least 5 eggs, with each lead value represented as a dot superimposed on a
standard boxplot showing median and quartile values

```{r variance_pb_5plus_eggs}
cv_label <- function(x) {
  return(data.frame(y = max(x) + 0.5, 
                    label = paste0(round(sd(x)/mean(x)*100, 1), "%")))
}

metal_subset %>% 
  add_count(premise_id) %>% 
  filter(n>4) %>% 
  filter(!str_detect(tolower(premise_id), pattern = "c")) %>% 
  mutate(premise_id = as.integer(premise_id)) %>% 
  arrange(premise_id) %>% 
  ggplot(mapping = aes(x = factor(premise_id), y = pb_ug_57g)) +
    geom_boxplot() +
    geom_dotplot(binaxis='y', 
                stackdir='center', 
                dotsize = 5, 
                binwidth = .02,
                fill="red") +
    stat_summary(fun.data = "cv_label", geom = "text", colour = "blue", size = 2) +
    stat_summary(fun.y = mean, geom = "point", colour = "purple", size = 1.5) +
    labs(x = "Premise ID", y = "ug Pb per 57g egg", 
         title = "Lead values for noncommercial premises with at least 5 eggs submitted (blue = CV%, purple = mean)") +
    theme(title=element_text(size=8,face="bold"), plot.title = element_text(hjust = 0.5))
    
    
```

Ideally the CV should be less than 20% (if replicates of the **same** sample then
less than 10%), so it looks like a large number of the submissions
are exhibiting high variability. Practically if all the values are less than 1ug it really is inconsequential - it's only a concern when the values approach level of concern. Variability in high valued samples has implications for pooling and how we 
should evaluate the non-pooled vs. pooled data. So far with our unpooled sample we've been
reporting maximum values, which given the high variability will make for a difficult
comparison to our pooled values (and awkward justification for pooling, apart from
our lack of funds).  We may need to backtrack and report mean values for the unpooled
samples to facilitate comparison across the two groups.
  
Here is the identical analysis for premises with two to four eggs:  

```{r variance_pb_2to4_eggs}
metal_subset %>% 
  add_count(premise_id) %>% 
  filter(n>1, n<5) %>% 
  filter(!str_detect(tolower(premise_id), pattern = "c")) %>% 
  mutate(premise_id = as.integer(premise_id)) %>% 
  arrange(premise_id) %>% 
  ggplot(mapping = aes(x = factor(premise_id), y = pb_ug_57g)) +
    geom_boxplot() +
    geom_dotplot(binaxis='y', 
                stackdir='center', 
                dotsize = 3, 
                binwidth = .02,
                fill="red") +
    stat_summary(fun.data = "cv_label", geom = "text", colour = "blue", size = 2) +
    stat_summary(fun.y = mean, geom = "point", colour = "purple", size = 1.5) +
    labs(x = "Premise ID", y = "ug Pb per 57g egg", 
         title = "Lead values for noncommercial premises with 2-4 eggs submitted (blue = CV%, purple = mean)") +
    theme(title = element_text(size=8,face="bold"), 
          plot.title = element_text(hjust = 0.5))
    
```

Cadmium values appear similar.  We have far fewer premises in this graph as
may of the cadmium values were below threshold.

```{r variance_cd_2plus_eggs}
metal_subset %>% 
  filter(!str_detect(tolower(premise_id), pattern = "c")) %>% 
  filter(!is.na(cd_ug_57g)) %>% 
  add_count(premise_id) %>% 
  filter(n > 1) %>% 
  mutate(premise_id = as.integer(premise_id)) %>% 
  arrange(premise_id) %>% 
  ggplot(mapping = aes(x = factor(premise_id), y = cd_ug_57g)) +
  geom_boxplot() +
  geom_dotplot(binaxis='y', 
               stackdir='center', 
               dotsize = 2, 
               binwidth = .02,
               fill="red") +
  stat_summary(fun.data = "cv_label", geom = "text", colour = "blue", size = 2) +
  stat_summary(fun.y = mean, geom = "point", colour = "purple", size = 1.5) +
  labs(x = "Premise ID", y = "ug Cd per 57g egg", 
       title = "Cadmium values for noncommercial premises with at least 2 eggs submitted (blue = CV%, purple = mean)") +
  theme(title = element_text(size=8,face="bold"), 
        plot.title = element_text(hjust = 0.5))
```

